---
title: "Ex3_bx2168_hl3339_wf2255"
author: "Banruo Xie;Wen Fan;Hanjun Li"
date: "3/7/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = F}
library(dplyr)
library(lubridate)
```

## (1)


$a_1$ means the probability of rainy day given the previous day is rainy day

$a_2$ means the probability of no rain day given the previous day is rainy day

$a_3$ means the probability of rainy day given the previous day is no rain day

$a_4$ means the probability of no rain day given the previous day is no rain day

## (2)
Let $X_i$ represent whether the ith day is rainy.

By Bayesian formula:

$P(X_n=0)=$
$P(X_n=0|X_{n-1}=0)P(X_{n-1}=0)+P(X_n=0|X_{n-1}=1)P(X_{n-1}=1)$
$= a_1P(X_n=0)+a_3(1-P(X_n=0))$

Therefore, $P(X_n=0)=\frac{a_3}{1-a_1+a_3}$

```{r}
data <- read.csv('./CentralPark.csv', header = T)
data$DATE <- as.POSIXct(strptime(as.character(data$DATE), "%m/%d/%y"))
data <- data %>% mutate(is_rain = if_else(PRCP>=1.5,T,F))
data$month <- month(data$DATE)
data$will_rain <- append(data$is_rain,c(NA))[2:(length(data$is_rain)+1)]
print(c((nrow(data %>% filter(month == 7, is_rain, will_rain)))/
          nrow(data %>% filter(month == 7, is_rain)),
        (nrow(data %>% filter(month == 7, is_rain, !will_rain)))/
          nrow(data %>% filter(month == 7, is_rain)),
        (nrow(data %>% filter(month == 7, !is_rain, will_rain)))/
          nrow(data %>% filter(month == 7, !is_rain)),
        (nrow(data %>% filter(month == 7, !is_rain, !will_rain)))/
          nrow(data %>% filter(month == 7, !is_rain))))
```

## (4)

Hypothesis test: $H_0: p_{00} = p_{11}$, $H_1:p_{00} \neq p_{11}$

$p_{00}$ is the probability of rainy day given the previous day is rainy day

$p_{11}$ is the probability of no rain day given the previous day is no rain day

Since $p_{00}$ and $p_{11}$ are independent, therefore, 
$\hat p_{00}\xrightarrow[\infty]{D}N(\hat p_{00},\frac{\hat p_{00}\left(1-\hat p_{00}\right)}{n_0})$

$\hat p_{11}\xrightarrow[\infty]{D}N(\hat p_{11},\frac{\hat p_{11}\left(1-\hat p_{11}\right)}{n_1})$

$\hat p_{00}-\hat p_{11}\xrightarrow[\infty]{D}N(0,\frac{\hat p_{00}\left(1-\hat p_{00}\right)}{n_0}-\frac{\hat p_{11}\left(1-\hat p_{11}\right)}{n_1})$

```{r}
a1 = (nrow(data %>% filter(month == 7, is_rain, will_rain)))/
  nrow(data %>% filter(month == 7, is_rain))
a4 = (nrow(data %>% filter(month == 7, !is_rain, !will_rain)))/
  nrow(data %>% filter(month == 7, !is_rain))
print(pnorm((a1-a4)/sqrt(a1*(1-a1)/nrow(data %>% filter(month == 7, is_rain))
                         +a4*(1-a4)/nrow(data %>% filter(month == 7, !is_rain)))))
```
Therefore, we reject $H_0$

## (5)

```{r}
data$will_rain2 <- append(data$will_rain,c(NA))[2:(length(data$will_rain)+1)]
```

$H_0$: Higher model chain can not improve.
$H_1$: Higher model chain does improve.

Using likelihood ratio test:

$\begin{aligned}\Lambda_{n} &=2\left\{\ell(\hat{\mathbf{P}})_{\text {second order}}-\ell(\hat{\mathbf{P}})_{\text {first order}}\right\}=2\left\{\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{r s t}-\sum_{s=1}^{S} \sum_{t=1}^{S}n_{. st} \log \hat{p}_{st}\right\} \\&=2\left\{\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{r s t}-\sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \hat{p}_{st}\right\}=2 \sum_{r=1}^{S} \sum_{s=1}^{S} \sum_{t=1}^{S} n_{r s t} \log \left(\frac{\hat{p}_{r s t}}{\hat{p}_{s t}}\right)\end{aligned}$

By asymptotic theory, 
$\Lambda_{n} \frac{\mathcal{D}}{n \rightarrow \infty} \chi_{(S-1)^{2}}^{2}$ 
```{r}
p00 <- (nrow(data %>% filter(month == 7, is_rain, will_rain)))/
  nrow(data %>% filter(month == 7, is_rain))
p01 <- (nrow(data %>% filter(month == 7, is_rain, !will_rain)))/
  nrow(data %>% filter(month == 7, is_rain))
p10 <- (nrow(data %>% filter(month == 7, !is_rain, will_rain)))/
  nrow(data %>% filter(month == 7, !is_rain))
p11 <- (nrow(data %>% filter(month == 7, !is_rain, !will_rain)))/
  nrow(data %>% filter(month == 7, !is_rain))

r000 <- nrow(data %>% filter(month == 7, is_rain, will_rain, will_rain2))
r001 <- nrow(data %>% filter(month == 7, is_rain, will_rain, !will_rain2))
r010 <- nrow(data %>% filter(month == 7, is_rain, !will_rain, will_rain2))
r011 <- nrow(data %>% filter(month == 7, is_rain, !will_rain, !will_rain2))
r100 <- nrow(data %>% filter(month == 7, !is_rain, will_rain, will_rain2))
r101 <- nrow(data %>% filter(month == 7, !is_rain, will_rain, !will_rain2))
r110 <- nrow(data %>% filter(month == 7, !is_rain, !will_rain, will_rain2))
r111 <- nrow(data %>% filter(month == 7, !is_rain, !will_rain, !will_rain2))

p000 <- r000/(r000 + r001)
p001 <- r001/(r000 + r001)
p010 <- r010/(r010 + r011)
p011 <- r011/(r010 + r011)
p100 <- r100/(r100 + r101)
p101 <- r101/(r100 + r101)
p110 <- r110/(r110 + r111)
p111 <- r111/(r110 + r111)

result <- 2* (r000*log(p000/p00) + r001*log(p001/p01) + r010*log(p010/p10)
              + r011*log(p011/p11) + r100*log(p100/p00) 
              + r101*log(p101/p01) + r110*log(p110/p10) 
              + r111*log(p111/p11)) 


pchisq(result,2)
```
 Therefore, we fail to reject $H_0$, higher model chain does not improve fit of the data.
