---
title: "HW2 Exercise 5"
author: "Wen Fan(wf2255), Banruo Xie(bx2168), Hanjun Li(hl3339)"
date: "3/7/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r, warning=FALSE}
library(readr)
library(dplyr)
transplant <- read_table2("~/Documents/Columbia/STAT W5703/HW/HW2/transplant.txt", 
    col_names = c("time", "type", "Indicator"), skip = 8)

# preprocessing
transplant$type <- factor(transplant$type)
type1_index <- which(transplant$type==1)
type2_index <- which(transplant$type==2)
death_index <- which(transplant$Indicator==1)
```

### Problem 1
Suppose the relapse rates for both treatment group are the same. i.e. $P_A = P_B = p$. We also see that the total number of death $n_d^{(k)}$ is independent of the number of at risk groups $n_A^{(k)}$ and $n_B^{(k)}$, so 
\begin{align*}
P(y^{(k)}=m|n_A^{(k)}=m_A,n_B^{(k)}=m_B,n_d^{(k)}=m_d) &= \frac{P(y^{(k)}=m, n_A^{(k)}=m_A,n_B^{(k)}=m_B,n_d^{(k)}=m_d)}{P(n_A^{(k)}=m_A,n_B^{(k)}=m_B,n_d^{(k)}=m_d)}\\
&= \frac{P(y^{(k)}=m,n_d^{(k)}=m_d| n_A^{(k)}=m_A,n_B^{(k)}=m_B)P(n_A^{(k)}=m_A,n_B^{(k)}=m_B)}{P(n_d^{(k)}=m_d| P(n_A^{(k)}=m_A,n_B^{(k)}=m_B)}\\
&= \frac{\binom{m_A}{m} p^m(1-p)^{m_A-m}\binom{m_B}{m_d-m} p^{m_d-m}(1-p)^{m_B-m_d+m}}{\binom{m_A+m_B}{m_d}p^{m_d}(1-p)^{m_A+m_B-m_d}}\\
&= \frac{\binom{m_A}{m}\binom{m_B}{m_d-m}}{\binom{m_A+m_B}{m_d}}
\end{align*}
Hence, it is a $HyperGeometric(n_{A}^{(k)}+n_{B}^{(k)}, n_{A}^{(k)}, n_{d}^{(k)})$

### Problem 2
From previous problem, we have the conditonal probability $$P(y^{(k)}=m|n_A^{(k)}=m_A,n_B^{(k)}=m_B,n_d^{(k)}=m_d)=\frac{\binom{n_A^{(k)}}{y^{(k)}}\binom{n_B^{(k)}}{n_d^{(k)}-y^{(k)}}}{\binom{n^{(k)}}{n_d^{(k)}}}$$
Using the formulas for the expected value and variance of a hypergeometric distribution: given $h(x; N,n,k)$, $$E[X]=\frac{n*k}{N}$$ $$Var(X)=\frac{n*k*(N-k)*(N-n)}{N^2*(N-1)}$$

so $$E^{(k)}=\frac{n_A^{(k)}n_d^{(k)}}{n^{(k)}}$$ $$V^{(k)}=\frac{n_A^{(k)}n_d^{(k)}(n^{(k)}-n_A^{(k)})(n^{(k)}-n_d^{(k)})}{(n^{(k)})^2(n^{(k)}-1)}=\frac{n_A^{(k)}n_d^{(k)}n_B^{(k)}n_s^{(k)}}{(n^{(k)})^2(n^{(k)}-1)}$$ 
Hence, shown.

### Problem 3
We have 
\begin{align*}
var(y^{(k)}-E^{(k)}) &= var(E[y^{(k)}-E^{(k)}| n_{A}^{(k)}, n_{B}^{(k)}, n_{d}^{(k)}])+E[var(y^{(k)}-E^{(k)}|n_{A}^{(k)}, n_{B}^{(k)}, n_{d}^{(k)})]\\
&= var(E[y^{(k)}|n_{A}^{(k)}, n_{B}^{(k)}, n_{d}^{(k)}]-E^{(k)})+E[V^{(k)}]\\
&= 0 + E[V^{(k)}]\\
&= E[V^{(k)}]
\end{align*}

### Problem 4
We have $$var[\sum_{k=1}^K(y^{k}-E^{(k)})]=\sum_{k=1}^{K} var[y^{(k)}-E^{(k)}]+2\sum_{i=1}^{K-1}\sum_{j=i+1}^{K}cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}]$$
Let the condition $(n_{A}^{(i)}, n_{B}^{(i)}, n_{d}^{(i)},n_{A}^{(j)}, n_{B}^{(j)}, n_{d}^{(j)})$ be $C$.
By the Law of Total Variance, 
\begin{align*}
cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}] &= cov[E[y^{(i)}-E^{(i)}| C],E[y^{(j)}-E^{(j)}|C]]+E[cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}|C]]\\
\end{align*}
We have $E[y^{(i)}-E^{(i)}|C]=E[y^{(j)}-E^{(j)}| C]=0$ and $cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}| C]=cov[y^{(i)},y^{(j)}| C]$ from problem 3, and $y^{(i)}$, $y^{(j)}$ are two independent hypergeometric variables; hence, $$cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}| C]=cov[y^{(i)},y^{(j)}| C]=0$$. As a result,
\begin{align*}
cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}] &= cov[E[y^{(i)}-E^{(i)}| C],E[y^{(j)}-E^{(j)}|C]]+E[cov[y^{(i)}-E^{(i)},y^{(j)}-E^{(j)}|C]]\\
&= 0+0 = 0
\end{align*}
and $$var[\sum_{k=1}^K(y^{k}-E^{(k)})]=\sum_{k=1}^{K} var[y^{(k)}-E^{(k)}]$$

### Problem 5
In our dataset, for each k, we have 1 death if the data is not censored, so $n_d^{k}=1$, $\forall k$
```{r}
n <- nrow(transplant)
k <- length(death_index)
na <- length(type1_index)
nb <- length(type2_index)
nd <- 1
df <- transplant[order(transplant$time),] %>% filter(Indicator==1)
y = numeric(k)
E = numeric(k)
Var = numeric(k)
for (i in seq_len(k)){
  E[i] <- na*nd/(na+nb)
  Var[i] <- na*nb*nd*(na+nb-nd)/((na+nb)^2*(na+nb-1))
  if(df[i,]$type==1){
    y[i] <- 1
    na <- na - 1
  } else {
    y[i] <- 0
    nb <- nb - 1
  }
}

Z <- sum(y-E)/sqrt(sum(Var))
pnorm(Z)
```
Using the logrank_test from "coin" library, we got the same conclusion:
```{r, message=FALSE}
library(coin)
logrank_test(Surv(time, Indicator)~type, data=transplant, distribution = "asymptotic")
```


Since the resulting p-value is insignificant, we do not have enough evidence to reject $H_0$.


